---
title: "Capstone - Toronto DineSafe Inspections"
author: "David Horan"
date: "November 8, 2016"
output: html_document
---

Step 1: Load the data from the DineSafe inspection database. The data was sourced in XML format.

```{r}
## Create a dataframe from the XML data.
library("XML")
dinesafe.tmp <- "C:/Users/Jenn & Dave/Documents/Capstone/dinesafe.Sept17.xml"
dinesafe <- xmlTreeParse(dinesafe.tmp)
class(dinesafe)
xmltop <- xmlRoot(dinesafe)
inspections <- xmlSApply(xmltop, function(x) xmlSApply(x, xmlValue))
inspect_df <- data.frame(t(inspections),row.names=NULL)

```

Step 2: Label the columns and set them to the appropriate data types.

```{r}
## Organize inspection data.
colnames(inspect_df)
inspect_df$ROW_ID <- as.numeric(inspect_df$ROW_ID)
inspect_df$ESTABLISHMENT_ID <- as.numeric(inspect_df$ESTABLISHMENT_ID)
inspect_df$ESTABLISHMENT_ID <- as.factor(inspect_df$ESTABLISHMENT_ID)
inspect_df$INSPECTION_ID <- as.numeric(inspect_df$INSPECTION_ID)
inspect_df$ESTABLISHMENT_NAME <- as.character(inspect_df$ESTABLISHMENT_NAME)
inspect_df$ESTABLISHMENTTYPE <- as.character(inspect_df$ESTABLISHMENTTYPE)
inspect_df$ESTABLISHMENTTYPE <- as.factor(inspect_df$ESTABLISHMENTTYPE)
inspect_df$ESTABLISHMENT_ADDRESS <- as.character(inspect_df$ESTABLISHMENT_ADDRESS)
inspect_df$ESTABLISHMENT_ADDRESS <- as.factor(inspect_df$ESTABLISHMENT_ADDRESS)
inspect_df$ESTABLISHMENT_STATUS <- as.character(inspect_df$ESTABLISHMENT_STATUS)
inspect_df$ESTABLISHMENT_STATUS <- as.factor(inspect_df$ESTABLISHMENT_STATUS)
inspect_df$MINIMUM_INSPECTIONS_PERYEAR <- as.numeric(inspect_df$MINIMUM_INSPECTIONS_PERYEAR)
inspect_df$MINIMUM_INSPECTIONS_PERYEAR <- as.factor(inspect_df$MINIMUM_INSPECTIONS_PERYEAR)
inspect_df$INFRACTION_DETAILS <- as.character(inspect_df$INFRACTION_DETAILS)
inspect_df$INFRACTION_DETAILS <- as.factor(inspect_df$INFRACTION_DETAILS)
inspect_df$INSPECTION_DATE <- as.character(inspect_df$INSPECTION_DATE)
inspect_df$INSPECTION_DATE <- as.Date(inspect_df$INSPECTION_DATE)
inspect_df$SEVERITY <- as.character(inspect_df$SEVERITY)
inspect_df$SEVERITY <- as.factor(inspect_df$SEVERITY)
inspect_df$ACTION <- as.character(inspect_df$ACTION)
inspect_df$ACTION <- as.factor(inspect_df$ACTION)
inspect_df$COURT_OUTCOME <- as.character(inspect_df$COURT_OUTCOME)
inspect_df$COURT_OUTCOME <- as.factor(inspect_df$COURT_OUTCOME)
inspect_df$AMOUNT_FINED <- as.numeric(inspect_df$AMOUNT_FINED)

## View a sample of the data.
head(inspect_df)

```

Step 3: Create a list of all unique inspection IDs.

```{r}
sub_insp <- subset(inspect_df,select = c(ESTABLISHMENT_ID:MINIMUM_INSPECTIONS_PERYEAR,INSPECTION_DATE))
inspect_unique <- unique(sub_insp)

head(inspect_unique)

```

Step 4: Create a list of inspection IDs that resulted in either Significant (S) or Crucial (C) severity.

```{r}
inspect_SevCru <- inspect_df[inspect_df$SEVERITY %in% c("C - Crucial","S - Significant"),3]

## Flag each of the unique inspections based on whether it resulted in a Significant or Crucial violation.
Sev_Cru <- vector()
for (i in 1:length(inspect_unique$INSPECTION_ID)){
  if (inspect_unique$INSPECTION_ID[i] %in% inspect_SevCru){
    Sev_Cru[i] <- T}
  else{
    Sev_Cru[i] <- F }}

## Bind the unique inspection records with the Severity indicator.
inspect_work <- as.data.frame(cbind(inspect_unique, Sev_Cru))

head(inspect_work)

```

We now have a list of unique inspections, flagged for whether or not they resulted in a Crucial or Significant violation.

Step 5: Add another potential explanatory variable: Month. This is a plausible variable, because there may be seasonal factors driving the inspection results.

```{r}
inspect_work$MONTH <- months(inspect_work$INSPECTION_DATE)
```

Step 6a: Check for complete cases:

```{r}
cc_test <- complete.cases(inspect_work)
length(which(cc_test == F))
```

Result: It appears that there are no incomplete cases in the remaining data.


Step 6b: Preliminary data analysis: 

```{r}
## Percentage of inspections resulting in a Significant or Crucial Health Violation.
nrow(inspect_work[Sev_Cru == 1,])/nrow(inspect_work)

```

Result: Approximately 22% of inspections resulted in a Significant or Crucial Health Violation.


Step 6c: Next, we will visualize the percentage of significant / crucial violations across the following categorical variables:

```{r}
## (i) Establishment Type
barplot(sort(table(inspect_work[Sev_Cru == 1,]$ESTABLISHMENTTYPE)/table(inspect_work$ESTABLISHMENTTYPE), decreasing = F), cex.names = 0.8)
## (ii) Minimum Inspections per Year
barplot(table(inspect_work[Sev_Cru == 1,]$MINIMUM_INSPECTIONS_PERYEAR)/table(inspect_work$MINIMUM_INSPECTIONS_PERYEAR))
## (iii) Month of Inspection
barplot(table(inspect_work[Sev_Cru == 1,]$MONTH)/table(inspect_work$MONTH), cex.names = 0.8)

```

Results:
(i) The appears to be some disparity in the rate of violations across Establishment Type.
(ii) Interestingly, the establisments that require 3 inspections per year also showed the highest incidence of Significant and Critical health violations. This makes intuitive sense, given that these are typically higher risk establishments.
(iii) There does appear to be some variation in the incidence of violations by month.


Step 7: Begin creating a logistic regression to evaluate the explanatory value of different variables on Significant and Crucial violations.

```{r}
glm_insp <- glm(Sev_Cru~ESTABLISHMENTTYPE+MINIMUM_INSPECTIONS_PERYEAR+MONTH,data = inspect_work, family = binomial("logit"))

summary(glm_insp)

```

